Formation of $\cenn$ Molecules in Oxygen-rich Interiors of Type II Supernovae

### 2013-01-14 doc/ms.tex
Add tables and other edits.

\newcommand{\ctwo}{{\rm C}_2}
\newcommand{\cthree}{{\rm C}_3}
\newcommand{\cfour}{{\rm C}_4}
\newcommand{\cfive}{{\rm C}_5}
\newcommand{\csix}{{\rm C}_6}
\newcommand{\cseven}{{\rm C}_7}
\newcommand{\ceight}{{\rm C}_8}
\newcommand{\ceightr}{{\rm C}_8^r}
\newcommand{\cenn}{{\rm C}_n}
\newcommand{\cotoco}{${\rm C} + {\rm O} \to {\rm CO} + \gamma$}
\newcommand{\twoctoctwo}{${\rm C} + {\rm C} \to \ctwo + \gamma$}
\newcommand{\coctoctwo}{${\rm CO} + {\rm C} \to \ctwo + {\rm O}$}
\newcommand{\ctowotococ}{$\ctwo + {\rm O} \to {\rm CO} + {\rm C}$}
\newcommand{\nar}{New Astron. Rev.}

@@

\author{Tianhong Yu, Donald D. Clayton, and Bradley S. Meyer}
\affil{Department of Physics and Astronomy, Clemson Universit, Clemson, SC 29634-0978}

@@

\keywords{globular clusters: general --- globular clusters: individual(NGC 6397,
NGC 6624, NGC 7078, Terzan 8}

@@

In basic agreement with this chemical model, \citet{2009ApJ...703..642C} raised
questions about the most effective way of producing $\ctwo$\ molecules,
which is the
first step to linear carbon chains. Using a large network of chemical
reactions, \citet{2009ApJ...703..642C} discovered that the neutral-neutral
reaction \coctoctwo
has a reasonably large cross section at temperatures near
5000K despite being endothermic by 4.8eV. Although the CO target abundance for
production of $\ctwo$\ in this manner must be calculated, they stated that at
$T=5000$K the rate factor
$\langle \sigma v \rangle = k_{CO,C} = 8.6 \times 10^{-14}$ cm$^3$ s$^{-1}$
created $\ctwo$\ much more
rapidly than the slow radiative reaction \twoctoctwo.
Sensing its significance,
they claimed in their list of conclusions: ``2. A new pathway to the formation
of carbon chains is active in the O-rich mass zone of the unmixed ejecta and
is identified as the CO conversion to $\ctwo$\ via collisions with C.''  They were
unable to evaluate its full quantitative consequences because their reaction
network terminated at $\cthree$\
and did not include the formation of linear carbon
chains $\cenn$
and their isomerization to carbon rings. Doing so is one goal of our
study. But the overriding goal is to point out large abundance differences
between the computations of these two groups and to resolve them if possible.

@@

Crucial to this task is the lifetime τCO of CO molecules. Their thermal
dissociation lifetime τγ is calculated from detailed balance with the
radiative association reaction, as in Section 2.1 of
\citet{2001ApJ...562..480C}.
Because of the large 11.1eV binding energy of the CO molecule,
$\tau_\gamma$ is very
temperature sensitive.  For readers numerical ease we will tabulate
$\tau_\gamma$ at
selected key temperatures in Table \ref{tab:quantities} to follow.
The flux and spectrum of newly
injected Compton electrons \citep{1991ApJ...375..221C} yields the lifetime
$\tau_e$ of CO
molecules against inelastic scattering dissociation by Compton electrons
caused by the 56Co radioactivity. In a gas of pure CO the mean energy per ion
pair is defined as the energy of primary electrons divided by the number of
pairs produced. \citet{1994ApJ...435..909L}
calculated the mean energy per ion pair
in pure CO gas and obtained the result $\Delta E = 32.3$ eV
deposited per dissociated CO
pair, which agrees well with the measurement of 32.2 eV by
\citet{1968...Klots}.
So the efficiency of energetic electrons for dissociating CO seems well
established. A lifetime $\tau_{CO}$ near one week is typical in SN 1987A, but
would be longer in SNII synthesizing less $^{56}$Co and at times greater than
about 8 months.

@@

Because our goal is to study the chemistry, we can take a very simple physical
model of the expansion; namely, temperature T=3800/(t/100d)=3.30x1010Ks/t(s),
where t is the time elapsed since core collapse and nO=1010cm-3 and
$n_C = 10^9$ cm$^{-3}$
at the starting time t=t(6000K) for our chemical network at
T=6000K. From our model choice for T(t) we get
$t(6000K) = 5.47 \times 10^6$s. At
subsequent times nO(t)=1010cm-3(t/t(6000))-3 owing to homologous expansion.
Let N be the number in a comoving, expanding, initially 1cm3 volume at 6000K
that held initially nO=1010cm-3 and nC=109cm-3. The only change of N during
expansion occurs through chemical reactions. N may be expressed as atom
fraction Y of the initial total number NO+NC= 1.10x1010 atoms. We
intentionally choose an O-rich interior having NO/NC=10, so that no carbon at
all would be able to condense if that interior were governed by chemical
equilibrium.

We use an intentionally limited set of chemical species because our goal is
to study carbon chemistry: C, O, CO, $\ctwo$, $\cthree$, ... $\ceight$\ and
$\ceightr$, the ringed isomer of $\ceight$, to which we give a lifetime
$\tau_8^r = 10$ s against thermal isomeric transition
from linear $\ceight$ to ringed $\ceightr$.
Those ringed molecules are the seeds for carbon
growth because their oxidation rates are much smaller than oxidation rates of
linear $\cenn$
whereas its C capture rates are fast. We take ringed $\ceightr$ to be
indestructible, simply integrating their rate of production. Our strategy is
to compute the number $\ceightr$ remaining after expansion from two differing
sources of $\ctwo$\ initiating the linear $\cenn$
chains. Those two source reactions are
\begin{itemize}
\item \twoctoctwo\ \citep{1999Sci...283.1290C}.
\item \coctoctwo\ \citep{2009ApJ...703..642C,2010ApJ...713....1C}.
\end{itemize}
We take our reaction rates from the rate tables in
\citet{2009ApJ...703..642C,2010ApJ...713....1C}.
The rate for the second reaction is indeed
much greater than that of the first reaction above T=3000K;
but it becomes increasingly the smaller of the two below T=2500K.
We include the thermal inverse reaction of every reaction,
which we calculate as in Eq.(3) of \citet{2001ApJ...562..480C}.
See Table 1 below for sample thermal dissociation rates for CO molecules.
We also include in our network dissociation of CO by Compton electrons.
CO is the only molecule for which thermal photodissociation is the dominant
destruction mechanism. For that rate we use
$\tau_e = 10^5$s $\exp( (t-10^6{\rm s}) / 111 {\rm d} )$,
not fit to any specific model but with plausibility for SN1987A.
It lengthens to $\tau_e = 10^6$ s near eight months.
Because of their helpful cataloging of reaction rate tables,
and in order to compare results without rate differences, we take the
rates as given in the tables of
\citet{2009ApJ...703..642C,2010ApJ...713....1C}.
As one example, we take the rate for \twoctoctwo\
as the rate given by RA4 in Table 5 in \citet{2009ApJ...703..642C}, p. 649.

Our computational reaction network is {\em libnucnet}
\citep{2007M&PSA..42.5215M} modified to carry the chemical
rather than nuclear network.  This network code has been thoroughly tested
on a wide variety of reaction networks and problems.
We scrupulously tested our network answers by hand calculations capable of
exposing coding errors.

@@

Evidently the abundance of CO within that zone during expansion attempts to
balance these creation and destruction effects, doing so exactly at the
time of maximum nCO. \citet{2013ApJ...762....5C}
has discussed the shape of the function
nCO(t). His Eq.(1) approximates the growth of nCO by its leading terms:
\begin{equation}
\frac{dn_{CO}}{dt} = n_{C(g)} n_O k_{CO} - \frac{n_{CO}}{\tau_{CO}} = 0
\label{eq:dncodt}
\end{equation}
at $t = t_{max}$.
The maximum abundance reached by CO is in cgs units
\begin{equation}
n_{CO}(t_max) = n_C n_O k_{CO} \tau_{CO} 
\label{eq:comax}
\end{equation}

@@

Figure \ref{fig:nco}
compares our network calculation of the abundance of CO molecules
with the expectation of Eq. (\ref{eq:comax}) when $\tau_\gamma$
dominates the destruction of CO.
Solid points display the equilibrium product nCOγeq calculated from the
factors shown in Table \ref{tab:quantities}
and from the number densities $n_C$ and $n_O$ after
their initial values at T=6000K have been reduced by the expansion factor
(t/t(6000K))-3. Tight agreement for $T > 3500$ K is immediately evident,
demonstrating that for $T > 3500$ K the abundance of CO is almost exactly in
thermal equilibrium. Below 3500K the abundance of CO becomes much less than
the expectation nCOγeq  of thermal equilibrium because the dissociation of CO
comes to be dominated by Compton electrons. The dashed vertical line at
T=3500K marks the approximate boundary between these two mechanisms for CO
dissociation. Notice carefully in Fig. \ref{fig:nco}
that nCO grows slowly, not reaching
its maximum until expansion has cooled to T=3500K.
At maximum nCO= 1.5x106cm-3, corresponding to YCO=0.7x10-3. The growth of nCO
shown in Fig. \ref{fig:nco} differs markedly from the results of
\citet{2009ApJ...703..642C}.
Their results (e.g. Fig. 11) show nCO climbing quickly near 6000K to a large
maximum number fraction near 0.1. That maximum would almost exhaust free
carbon. Our results in Fig. \ref{fig:nco}
so differ from theirs that the difference
must be resolved. \citet{2013ApJ...762....5C} has analyzed the expectation
of growth of nCO to a single maximum before declining, and our results are
in line with that expectation.

These features are further detailed in Fig. \ref{fig:flows},
which shows the reaction
currents (reactions per second per atom) into and out of CO. There exists a
near steady state in that the production of CO is almost balanced by the
two flows destroying CO. The destruction flow $n_{CO}/\tau_\gamma$
almost balances the
association flow above 4000K except for a tiny excess leading to the
slow growth of nCO evident in Fig. \ref{fig:nco}. In that temperature range the
production by \cotoco\ in equilibrium with the thermal radiation field,
as Fig. \ref{fig:nco} implied. The destruction flow $n_{CO}/\tau_e$
almost balances the
creation flow below 3500K. The transition between destruction modes occurs
near 3500K. It is no coincidence that the maximum of nCO occurs when
Compton electrons begin to dominate CO dissociation, because nCO would
continue growing as long as thermal photons dominate CO dissociation.

@@

Figure \ref{fig:cn} shows the shape of $\cenn$\ vs. $n$ for $\cthree$\ to
$\ceight$\ at two different temperatures near 2000K. As temperature declines,
the curve flattens because photoejection from $\cenn$\ by thermal photons
(or phonons) weakens. These abundance ratios are almost in a steady state,
but that steady state changes slightly as T falls. These patterns show
almost equal values for $Y(\cthree)$ because that curve is relatively flat
with time near $t-t(6000K) = 10^7$ s (Fig. \ref{fig:yi}).
Table \ref{tab:flows} lists our computed reaction flows both in and out of
$\csix$\ at T = 2031K and T = 1802K. The reactions are specified there in the
compact notation target(in,out)residual. Fig. \ref{fig:cn} and
Table \ref{tab:flows} are made to be studied together.
Note that the abundances of $\ceight$\ are very small:
$Y(C_8)= 10^{-36}$ at T = 1802K. The isomeric transition from linear to
ringed $\ceight$, for which we estimate a lifetime $\tau = 10$ s,
provides the nucleations for graphite growth. Figure \ref{fig:ncoc2} showed
that $Y(C_{8}^r)$

We call attention to several conclusions to be drawn from Table \ref{tab:flows}.
Firstly, the strongest flows by a wide margin involving C6 are
$C_5 + C \to C_6$,
which is very fast ($k = 3 \times 10^{-10} cm^3 s^{-1}$)
because vibrational excitation of $\csix$
obviates the need for creating a photon \citep{1999Sci...283.1290C},
and its inverse reaction which ejects a C atom. Furthermore, those flows
are equal to three significant figures, illustrating the near steady
state of the abundance pattern. Secondly, thermal dissociation of $\csix$\ is
very much faster than its oxidation, showing the small effect of oxidation
on the abundance pattern. The destroying flows from $\csix$\ to $\cfive$
stand in the approximate ratio $C_6(\gamma,C)/C_6(O,CO) = 5 \times 10^5$
at $T = 2031$ K and $2 \times 10^4$ at $T = 1802$ K.
Photodissociation dominates. The smaller value for that ratio at T=1802K
occurs because oxidation maintains its effectiveness as T drops but
thermal disruption does not, being much more temperature sensitive.
For this reason the abundance pattern is flatter and C6 is approximately
3000 times more abundant at T=1802K than at T=2031K.
It is for that reason that destruction flows in Table \ref{tab:flows}
for C6 are
larger at the smaller temperature.
The steady state shifts to new ratios as T falls.
Thirdly, calculation of the rates will be illustrated for sake of clarity
by the flow $C_6 + O \to CO + C_5$.
From Fig. \ref{fig:cn}
one sees that at T=2031K the abundance $Y(\csix) = 10^{-33.3}$.
The value of $Y(O) = 10^{10} cm^{-3} / 1.1 \times 10^{10} cm^{-3} = 0.909$,
so the flow per atom per
second (e.g. Eq.3) is dY/dt= Y(C6)Y(O)n(2031K)k(C6O).
Expansion from T=6000K to T=2031K has diluted the total number
density n to 1.1x1010cm-3 (2031K/6000K)3= 4.27x108cm-3.
The fast oxidation reaction rate is  k(C6O)=3x10-10cm3s-1.
Gathering factors yields approximately 4x10-35 n-1s-1,
in good approximation to the Table \ref{tab:flows}
flow entry $3.61 \times 10^{-35} n^{-1}s^{-1}$.

@@

\begin{figure}
\epsscale{.80}
%\plotone{f1.eps}
\caption{
Comparison of the density of CO molecules calculated by the numerical
network (curve) with the expectation nCOγeq = nCnO$k_{CO}$τγ for an abundance
nCO in thermal equilibrium (dots) with photons.
Above T=4000K the CO abundance is seen to be in an accurate thermal
equilibrium, but at lower temperature nCO is increasingly smaller than
thermal equilibrium would require. Instead of thermal photons,
the destruction of CO becomes increasingly dominated by
Compton electrons for $T < 3500$ K.
One see in Fig. \ref{fig:nco} that nCOmax occurs near 3500K,
which occurs near $10^7$ s (see Table \ref{tab:quantities})}. 
\label{fig:nco}
\end{figure}

@@

\begin{figure}
\epsscale{.80}
%\plotone{f1.eps}
\caption{
Flows into and out of CO illustrate the near steady state between creation
and destruction and the transition region near 3500K between the two modes
of destruction of CO. The destruction flow nCOnC$k_{CO}$+C is only $3.27 \times
10^{-16}$ atom$^{-1}$s$^{-1}$, too small to be visible in Fig. \ref{fig:flows}. 
} \label{fig:flows}
\end{figure}

@@

\begin{deluxetable}{ccrrrrrrrrcrl}
\tabletypesize{\scriptsize}
\rotate
\tablecaption{Values of $\tau_\gamma$(CO) and related quantities at
selected $T$}
\label{tab:quantities}
\tablewidth{0pt}
\tablehead{
\colhead{POS} & \colhead{chip} & \colhead{ID} & \colhead{X} & \colhead{Y} &
\colhead{RA} & \colhead{DEC} & \colhead{IAU$\pm$ $\delta$ IAU} &
\colhead{IAP1$\pm$ $\delta$ IAP1} & \colhead{IAP2 $\pm$ $\delta$ IAP2} &
\colhead{star} & \colhead{E} & \colhead{Comment}
}
\startdata
0 & 2 & 1 & 1370.99 & 57.35    &   6.651120 &  17.131149 & 21.344$\pm$0.006  & 2
4.385$\pm$0.016 & 23.528$\pm$0.013 & 0.0 & 9 & -    \\
0 & 2 & 2 & 1476.62 & 8.03     &   6.651480 &  17.129572 & 21.641$\pm$0.005  & 2
3.141$\pm$0.007 & 22.007$\pm$0.004 & 0.0 & 9 & -    \\
0 & 2 & 3 & 1079.62 & 28.92    &   6.652430 &  17.135000 & 23.953$\pm$0.030  & 2
4.890$\pm$0.023 & 24.240$\pm$0.023 & 0.0 & - & -    \\
0 & 2 & 4 & 114.58  & 21.22    &   6.655560 &  17.148020 & 23.801$\pm$0.025  & 2
5.039$\pm$0.026 & 24.112$\pm$0.021 & 0.0 & - & -    \\
0 & 2 & 5 & 46.78   & 19.46    &   6.655800 &  17.148932 & 23.012$\pm$0.012  & 2
3.924$\pm$0.012 & 23.282$\pm$0.011 & 0.0 & - & -    \\
0 & 2 & 6 & 1441.84 & 16.16    &   6.651480 &  17.130072 & 24.393$\pm$0.045  & 2
6.099$\pm$0.062 & 25.119$\pm$0.049 & 0.0 & - & -    \\
0 & 2 & 7 & 205.43  & 3.96     &   6.655520 &  17.146742 & 24.424$\pm$0.032  & 2
5.028$\pm$0.025 & 24.597$\pm$0.027 & 0.0 & - & -    \\
0 & 2 & 8 & 1321.63 & 9.76     &   6.651950 &  17.131672 & 22.189$\pm$0.011  & 2
4.743$\pm$0.021 & 23.298$\pm$0.011 & 0.0 & 4 & edge \\
\enddata
\end{deluxetable}

%% If you use the table environment, please indicate horizontal rules using
%% \tableline, not \hline.
%% Do not put multiple tabular environments within a single table.
%% The optional \label should appear inside the \caption command.

>>>>

\newcommand{\ctwo}{{\rm C}_2}
\newcommand{\cthree}{{\rm C}_3}
\newcommand{\cfour}{{\rm C}_4}
\newcommand{\cfive}{{\rm C}_5}
\newcommand{\csix}{{\rm C}_6}
\newcommand{\cseven}{{\rm C}_7}
\newcommand{\ceight}{{\rm C}_8}
\newcommand{\ceightr}{{\rm C}_8^r}
\newcommand{\cenn}{{\rm C}_n}
\newcommand{\cotoco}{${\rm C} + {\rm O} \to {\rm CO} + \gamma$}
\newcommand{\twoctoctwo}{${\rm C} + {\rm C} \to \ctwo + \gamma$}
\newcommand{\coctoctwo}{${\rm CO} + {\rm C} \to \ctwo + {\rm O}$}
\newcommand{\ctowotococ}{$\ctwo + {\rm O} \to {\rm CO} + {\rm C}$}
\newcommand{\cfivectocsix}{$\cfive + {\rm C} \to \csix$}
\newcommand{\csixotococfive}{${\rm C}_6 + {\rm O} \to {\rm CO} + {\rm C_5}$}
\newcommand{\ncogeq}{n_{CO}^{\gamma eq}}
\newcommand{\nar}{New Astron. Rev.}

@@

\author{Tianhong Yu, Bradley S. Meyer, and Donald D. Clayton}
\affil{Department of Physics and Astronomy, Clemson Universit, Clemson, SC 29634-0978}

@@

\keywords{atomic processes, ISM: abundances, ISM: general, meteorites, meteors, meteoroids, supernovae: general, supernovae: individual: SN 1987A}

@@

In basic agreement with this chemical model, \citet{2009ApJ...703..642C} raised
questions about the most effective way of producing $\ctwo$\ molecules,
which is the
first step to linear carbon chains. Using a large network of chemical
reactions, \citet{2009ApJ...703..642C} discovered that the neutral-neutral
reaction \coctoctwo\
has a reasonably large cross section at temperatures near
5000K despite being endothermic by 4.8eV. Although the CO target abundance for
production of $\ctwo$\ in this manner must be calculated, they stated that at
$T=5000$K the rate factor
$\langle \sigma v \rangle = k_{CO,C} = 8.6 \times 10^{-14}$ cm$^3$ s$^{-1}$
created $\ctwo$\ much more
rapidly than the slow radiative reaction \twoctoctwo.
Sensing its significance,
they claimed in their list of conclusions: ``2. A new pathway to the formation
of carbon chains is active in the O-rich mass zone of the unmixed ejecta and
is identified as the CO conversion to $\ctwo$\ via collisions with C.''  They were
unable to evaluate its full quantitative consequences because their reaction
network terminated at $\cthree$\
and did not include the formation of linear carbon
chains $\cenn$
and their isomerization to carbon rings. Doing so is one goal of our
study. But the overriding goal is to point out large abundance differences
between the computations of these two groups and to resolve them if possible.

@@

Crucial to this task is the lifetime $\tau_{CO}$ of CO molecules. Their thermal
dissociation lifetime $\tau_\gamma$
is calculated from detailed balance with the
radiative association reaction, as in Section 2.1 of
\citet{2001ApJ...562..480C}.
Because of the large 11.1eV binding energy of the CO molecule,
$\tau_\gamma$ is very
temperature sensitive.  For readers' numerical ease we will tabulate
$\tau_\gamma$ at
selected key temperatures in Table \ref{tab:quantities} to follow.
The flux and spectrum of newly
injected Compton electrons \citep{1991ApJ...375..221C} yields the lifetime
$\tau_e$ of CO
molecules against inelastic scattering dissociation by Compton electrons
caused by the $^{56}$Co
radioactivity. In a gas of pure CO the mean energy per ion
pair is defined as the energy of primary electrons divided by the number of
pairs produced. \citet{1994ApJ...435..909L}
calculated the mean energy per ion pair
in pure CO gas and obtained the result $\Delta E = 32.3$ eV
deposited per dissociated CO
pair, which agrees well with the measurement of 32.2 eV by
\citet{1968...Klots}.
So the efficiency of energetic electrons for dissociating CO seems well
established. A lifetime $\tau_{CO}$ near one week is typical in SN 1987A, but
would be longer in SNII synthesizing less $^{56}$Co and at times greater than
about 8 months.

@@

Because our goal is to study the chemistry, we can take a very simple physical
model of the expansion; namely, temperature T=3800/(t/100d)=3.30x1010Ks/t(s),
where t is the time elapsed since core collapse and nO=1010cm-3 and
$n_C = 10^9$ cm$^{-3}$
at the starting time t=t(6000K) for our chemical network at
T=6000K. From our model choice for T(t) we get
$t(6000K) = 5.47 \times 10^6$s. At
subsequent times $n_O(t) = 10^{10}{\rm cm}^{-3}(t/t(6000 {\rm K}))^{-3}$
owing to homologous expansion.
Let $N$ be the number in a comoving, expanding, initially 1 cm$^3$
volume at 6000 K
that held initially $n_O=10^{10}$ cm$^{-3}$ and $n_C = 10^9$ cm$^{-3}$.
The only change of $N$ during
expansion occurs through chemical reactions. $N$ may be expressed as atom
fraction $Y$ of the initial total number
$N_O + N_C = 1.10 \times 10^{10}$ atoms. We
intentionally choose an O-rich interior having $N_O/N_C = 10$,
so that no carbon at
all would be able to condense if that interior were governed by chemical
equilibrium.

We use an intentionally limited set of chemical species because our goal is
to study carbon chemistry: C, O, CO, $\ctwo$, $\cthree$, ... $\ceight$\ and
$\ceightr$, the ringed isomer of $\ceight$, to which we give a lifetime
$\tau_8^r = 10$ s against thermal isomeric transition
from linear $\ceight$ to ringed $\ceightr$.
Those ringed molecules are the seeds for carbon
growth because their oxidation rates are much smaller than oxidation rates of
linear $\cenn$
whereas its C capture rates are fast. We take ringed $\ceightr$ to be
indestructible, simply integrating their rate of production. Our strategy is
to compute the number $\ceightr$ remaining after expansion from two differing
sources of $\ctwo$\ initiating the linear $\cenn$
chains. Those two source reactions are
\begin{itemize}
\item \twoctoctwo\ \citep{1999Sci...283.1290C}.
\item \coctoctwo\ \citep{2009ApJ...703..642C,2010ApJ...713....1C}.
\end{itemize}
We take our reaction rates from the rate tables in
\citet{2009ApJ...703..642C,2010ApJ...713....1C}.
The rate for the second reaction is indeed
much greater than that of the first reaction above T=3000K;
but it becomes increasingly the smaller of the two below T=2500K.
We include the thermal inverse reaction of every reaction,
which we calculate as in Eq.(3) of \citet{2001ApJ...562..480C}.
See Table 1 below for sample thermal dissociation rates for CO molecules.
We also include in our network dissociation of CO by Compton electrons.
CO is the only molecule for which thermal photodissociation is the dominant
destruction mechanism. For that rate we use
$\tau_e (s) = 10^5 \exp( (t-10^6{\rm s}) / 111 {\rm d} )$,
not fit to any specific model but with plausibility for SN1987A.
It lengthens to $\tau_e = 10^6$ s near eight months.
Because of their helpful cataloging of reaction rate tables,
and in order to compare results without rate differences, we take the
rates as given in the tables of
\citet{2009ApJ...703..642C,2010ApJ...713....1C}.
As one example, we take the rate for \twoctoctwo\
as the rate given by RA4 in Table 5 in \citet{2009ApJ...703..642C}, p. 649.

Our computational reaction network is {\em libnucnet}
(\citealp{2007M&PSA..42.5215M}--see also
\url{http://sourceforge.net/projects/libnucnet/}) modified to follow
the chemical rather than nuclear reactions.
This network code has been thoroughly tested
on a wide variety of reaction networks and problems.
We scrupulously tested our network answers by hand calculations capable of
exposing coding errors.

@@

Evidently the abundance of CO within that zone during expansion attempts to
balance these creation and destruction effects, doing so exactly at the
time of maximum $n_{CO}$. \citet{2013ApJ...762....5C}
has discussed the shape of the function
$n_{CO}$(t). His Eq.(1) approximates the growth of $n_{CO}$ by its leading terms:
\begin{equation}
\frac{dn_{CO}}{dt} = n_{C(g)} n_O k_{CO} - \frac{n_{CO}}{\tau_{CO}} = 0
\label{eq:dncodt}
\end{equation}
at $t = t_{max}$.
The maximum abundance reached by CO is in cgs units
\begin{equation}
n_{CO}(t_{max}) = n_C n_O k_{CO} \tau_{CO} 
\label{eq:comax}
\end{equation}

@@

Figure \ref{fig:nco}
compares our network calculation of the abundance of CO molecules
with the expectation of Eq. (\ref{eq:comax}) when $\tau_\gamma$
dominates the destruction of CO.
Solid points display the equilibrium product $\ncogeq$ calculated from the
factors shown in Table \ref{tab:quantities}
and from the number densities $n_C$ and $n_O$ after
their initial values at T=6000K have been reduced by the expansion factor
(t/t(6000K))-3. Tight agreement for $T > 3500$ K is immediately evident,
demonstrating that for $T > 3500$ K the abundance of CO is almost exactly in
thermal equilibrium. Below 3500K the abundance of CO becomes much less than
the expectation $\ncogeq$ of thermal equilibrium because the dissociation of CO
comes to be dominated by Compton electrons. The dashed vertical line at
T=3500K marks the approximate boundary between these two mechanisms for CO
dissociation. Notice carefully in Fig. \ref{fig:nco}
that $n_{CO}$ grows slowly, not reaching
its maximum until expansion has cooled to T=3500K.
At maximum $n_{CO} = 1.5 \times 10^6$ cm$^{-3}$,
corresponding to $Y(CO) = 0.7 \times 10^{-3}$. The growth of $n_{CO}$
shown in Fig. \ref{fig:nco} differs markedly from the results of
\citet{2009ApJ...703..642C}.
Their results (e.g. Fig. 11) show $n_{CO}$ climbing quickly near 6000K to a large
maximum number fraction near 0.1. That maximum would almost exhaust free
carbon. Our results in Fig. \ref{fig:nco}
so differ from theirs that the difference
must be resolved. \citet{2013ApJ...762....5C} has analyzed the expectation
of growth of $n_{CO}$ to a single maximum before declining, and our results are
in line with that expectation.

These features are further detailed in Fig. \ref{fig:flows},
which shows the reaction
currents (reactions per second per atom) into and out of CO. There exists a
near steady state in that the production of CO is almost balanced by the
two flows destroying CO. The destruction flow $n_{CO}/\tau_\gamma$
almost balances the
association flow above 4000K except for a tiny excess leading to the
slow growth of $n_{CO}$ evident in Fig. \ref{fig:nco}. In that temperature range the
production by \cotoco\ in equilibrium with the thermal radiation field,
as Fig. \ref{fig:nco} implied. The destruction flow $n_{CO}/\tau_e$
almost balances the
creation flow below 3500K. The transition between destruction modes occurs
near 3500K. It is no coincidence that the maximum of $n_{CO}$ occurs when
Compton electrons begin to dominate CO dissociation, because $n_{CO}$ would
continue growing as long as thermal photons dominate CO dissociation.

@@

Figure \ref{fig:cn} shows the shape of $\cenn$\ vs. $n$ for $\cthree$\ to
$\ceight$\ at two different temperatures near 2000K. As temperature declines,
the curve flattens because photoejection from $\cenn$\ by thermal photons
(or phonons) weakens. These abundance ratios are almost in a steady state,
but that steady state changes slightly as T falls. These patterns show
almost equal values for $Y(\cthree)$ because that curve is relatively flat
with time near $t-t(6000K) = 10^7$ s (Fig. \ref{fig:yi}).
Table \ref{tab:flows} lists our computed reaction flows both in and out of
$\csix$\ at T = 2031K and T = 1802K. The reactions are specified there in the
compact notation target(in,out)residual. Fig. \ref{fig:cn} and
Table \ref{tab:flows} are made to be studied together.
Note that the abundances of $\ceight$\ are very small:
$Y(\ceight)= 10^{-36}$ at T = 1802K. The isomeric transition from linear to
ringed $\ceight$, for which we estimate a lifetime $\tau = 10$ s,
provides the nucleations for graphite growth. Figure \ref{fig:ncoc2} showed
that $Y(\ceightr)$

We call attention to several conclusions to be drawn from Table \ref{tab:flows}.
Firstly, the strongest flows by a wide margin involving $\csix$\ are
\cfivectocsix,
which is very fast ($k = 3 \times 10^{-10} cm^3 s^{-1}$)
because vibrational excitation of $\csix$
obviates the need for creating a photon \citep{1999Sci...283.1290C},
and its inverse reaction which ejects a C atom. Furthermore, those flows
are equal to three significant figures, illustrating the near steady
state of the abundance pattern. Secondly, thermal dissociation of $\csix$\ is
very much faster than its oxidation, showing the small effect of oxidation
on the abundance pattern. The destroying flows from $\csix$\ to $\cfive$
stand in the approximate ratio
${\rm C}_6(\gamma,{\rm C})/{\rm C_6}({\rm O},{\rm CO}) = 5 \times 10^5$
at $T = 2031$ K and $2 \times 10^4$ at $T = 1802$ K.
Photodissociation dominates. The smaller value for that ratio at T=1802K
occurs because oxidation maintains its effectiveness as T drops but
thermal disruption does not, being much more temperature sensitive.
For this reason the abundance pattern is flatter and $\csix$\ is approximately
3000 times more abundant at T=1802K than at T=2031K.
It is for that reason that destruction flows in Table \ref{tab:flows}
for $\csix$\ are
larger at the smaller temperature.
The steady state shifts to new ratios as T falls.
Thirdly, calculation of the rates will be illustrated for sake of clarity
by the flow
\csixotococfive.
From Fig. \ref{fig:cn}
one sees that at $T = 2031$ K the abundance $Y(\csix) = 10^{-33.3}$.
The value of $Y(O) = 10^{10} cm^{-3} / 1.1 \times 10^{10} cm^{-3} = 0.909$,
so the flow per atom per
second (e.g., Eq. \ref{eq:dycodt}) is $dY/dt= Y(\csix)Y(O)n(2031K)k(C6O)$.
Expansion from T=6000K to T=2031K has diluted the total number
density n to 1.1x1010cm-3 (2031K/6000K)3= 4.27x108cm-3.
The fast oxidation reaction rate is  k(C6O)=3x10-10cm3s-1.
Gathering factors yields approximately 4x10-35 n-1s-1,
in good approximation to the Table \ref{tab:flows}
flow entry $3.61 \times 10^{-35} n^{-1}s^{-1}$.

@@

\begin{figure}
\epsscale{.80}
%\plotone{f1.eps}
\caption{
Comparison of the density of CO molecules calculated by the numerical
network (curve) with the expectation $\ncogeq = n_C n_O k_{CO} \tau_\gamma$
for an abundance
$n_{CO}$ in thermal equilibrium (dots) with photons.
Above T=4000K the CO abundance is seen to be in an accurate thermal
equilibrium, but at lower temperature $n_{CO}$ is increasingly smaller than
thermal equilibrium would require. Instead of thermal photons,
the destruction of CO becomes increasingly dominated by
Compton electrons for $T < 3500$ K.
One see in Fig. \ref{fig:nco} that $n_{CO}$max occurs near 3500K,
which occurs near $10^7$ s (see Table \ref{tab:quantities})}. 
\label{fig:nco}
\end{figure}

@@

\begin{figure}
\epsscale{.80}
%\plotone{f1.eps}
\caption{
Flows into and out of CO illustrate the near steady state between creation
and destruction and the transition region near 3500K between the two modes
of destruction of CO. The destruction flow $n_{CO}$nC$k_{CO}$+C is only $3.27 \times
10^{-16}$ atom$^{-1}$s$^{-1}$, too small to be visible in Fig. \ref{fig:flows}. 
} \label{fig:flows}
\end{figure}

@@

\begin{table}
\begin{center}
\caption{Values of $\tau_\gamma$(CO) and related quantities at
selected time $t$ and temperature $T$}
\label{tab:quantities}
\begin{tabular}{ccccccc}
\tableline\tableline
$t$ (s) & $t - t(6000 {\rm K})$ (s) & $\left(t / t(6000 {\rm K} \right)^{-3}$ & $T$ (K) & $\tau_\gamma$ (s) & $k_{CO}$ (cm$^{-3}$ s$^{-1}$) & $\ncogeq$ (cm$^{-3}$) \\
\tableline
$5.47 \times 10^6$ &
  0 &
  1 &
  6000 &
  $4.22 \times 10^{-2}$ &
  $3.13 \times 10^{-17}$ &
  $1.32 \times 10^1$ \\
$5.97 \times 10^6$ &
  $4.97 \times 10^5$ &
  0.769 &
  5500 &
  $3.54 \times 10^{-1}$ &
  $2.99 \times 10^{-17}$ &
  $6.27 \times 10^1$ \\
$6.57 \times 10^6$ &
  $1.09 \times 10^6$ &
  0.578 &
  5000 &
  $4.46 \times 10^{0}$ &
  $2.83 \times 10^{-17}$ &
  $4.24 \times 10^2$ \\
$7.30 \times 10^6$ &
  $1.82 \times 10^6$ &
  0.422 &
  4500 &
  $9.70 \times 10^{1}$ &
  $2.67 \times 10^{-17}$ &
  $4.61 \times 10^3$ \\
$8.21 \times 10^6$ &
  $2.74 \times 10^6$ &
  0.296 &
  4000 &
  $4.43 \times 10^{3}$ &
  $2.48 \times 10^{-17}$ &
  $9.67 \times 10^4$ \\
$9.38 \times 10^6$ &
  $3.91 \times 10^6$ &
  0.198 &
  3500 &
  $5.86 \times 10^{5}$ &
  $2.28 \times 10^{-17}$ &
  $5.26 \times 10^6$ \\
$1.09 \times 10^7$ &
  $5.47 \times 10^6$ &
  0.125 &
  3000 &
  $3.77 \times 10^{8}$ &
  $2.05 \times 10^{-17}$ &
  $1.21 \times 10^9$ \\
\tableline
\end{tabular}
\end{center}
\end{table}
